\( \tool \): JavaScript IR-based Semantics Extraction Toolchain

### 2019-08-20 icse/impl.tex
Changed filename framework.tex -> impl.tex
Edited implementation section

\section{Building entire system to extract semantics}\label{sec:framework}

In this section, we demonstrate how our idea can be implemented for target language. Although it is hard to generalize the process because the format of specification or
environment of the programming language is varying, We shows the possibility to build entire system to extract target language specification and interpreting it with minimal human effort by giving end-to-end example of ECMAScript language.

\subsection{Extracting specification}
First step is to collect language specification and extract component that we can process automatically. For our target language, ECMAScript, has HTML format of specification in their website, and
in the specification they have BNF form of syntax, highly structured algorithm steps, and verbose texts to describe additional environment to evaluating ECMAScript source code.
We get these information by parsing HTML document, and passing these result to our previous module to get AST, Parser, and executable algorithms.

To organize these primitives and running the program, we have to model initial environment of language such as initialization of builtin types, or entry point of the program, etc.

\subsection{Generating initial environment for execution}
For ECMAScript Specification, initial environment consists of the Builtin objects, and Intrinsic objects. We describe how to generate each component.

\subsubsection{Builtin objects}

 Builtin object is a pre-defined object with many built-in functions. For example, In ECMAScript \( \code{Array} \) corresponds to an object which can generate an array by
 calling it. Each builtin object provides data property and functions, just as ordinary object. When its function is called,
 its algorithm specified in the document should be called.

 In specification, a builtin function object usually matches to abstract algorithm. Its algorithm name
 describes the hierarchy about object it should be contained. For example, there is an abstract algorithm named \( \code{ Object.prototype.propertyIsEnumerable } \). With its name,
 we can know that it corresponds to builtin function in the object \( \code{ Object.prototype } \) as property.
 To support builtin object, we make structure of builtin object and its properties.
 After we correctly organizes structure, then we made algorithm that automatically map corresponding field to the abstract algorithm in specification by matching its function name.

 \subsubsection{Intrinsic objects}

 In specification, "Well-known intrinsics are built-in objects that are explicitly referenced by the algorithms".
 Abstract algorithm uses identifiers like "\%name\%" to point builtin objects. For example it refers builtin Array object as \( \code{\%Array\%} \) .
 Map between Intrinsic name and Builtin objects are available in specification as table form in section "6.1.7.4 Well-Known Intrinsic Objects", so we parse this table to generate global identifier to object.

\subsection{Algorithm scope}

 In specification, All algorithm has name to refer it, but some algorithms have same names and they are called in context-dependent way.
In "5.2 Algorithm Conventions", it defines three types of abstract operation.
 \begin{itemize}
  \item Abstract operation: they can be called by name in any context
  \item Method-like abstract operation: algorithm dispatch depends on receiver object
  \item Syntax-directed operation: algorithm dispatch depends on receiver AST
  \end{itemize}

First, we categorized sections by types of algorithm.
Method-like abstract operation has separate sections to describe it, so we manually add these sections to treat differently.
Also Syntax-directed operations are described in distinct sections, so both case is easy to extract it.
Rest algorithms are considered as ordinary Abstract operation.

\subsubsection{Abstract operation}

We just convert abstract operations to function, and bind global identifier to that function to call it by algorithm name.

\subsubsection{Method-like abstract operation}

In ECMAScript specification, there are some objects which has similar interface, but they have different algorithms.

For Example, Ordinary Object and String Exotic Object is different object in specification, but they have same name of algorithm(i.e. \( \code{[[getOwnProperty]](N)} \))

In this case, algorithm dispatching should consider receiver Object(i.e. \( \code{x.[[GetOwnProperty]](N)} \)).
For method-like abstract operation, we organized map from algorithm name to function for each type,
 and we insert functions as property of object when such type of object is created. So In our example, when  \( \code{x} \) is created with Ordinary Object type,
 then we insert property with name \( \code{[[GetOwnProperty]]} \) and function from method-like abstract operation "[[GetOwnProperty]]" in Ordinary Object Internal method section, at the time of map creation.

 \subsubsection{Syntax-directed operation}

 Like method-like abstract operation, many AST objects have similar interface, but they have different algorithm.
The most frequent example is "Evaluation" algorithm of AST Values. Many non-terminal symbol have different algorithm with name "Evaluation", and
when specification requires "the result of evaluating \textit{variable}", then it should apply "Evaluation" of non-terminal symbol which matches to \textit{variable}.
It is assumed that \textit{variable} is a AST value. Also in Syntax-directed operation, the sub-component of AST can be used as values. For example, In the
Evaluation algorithm of \textit{CallExpression} : \textit{CallExpression} \textit{Arguments}, algorithm can uses \textit{CallExpression} or \textit{Arguments} to refer sub-component of AST.
 To support Syntax-directed operation, we insert converted functions in AST structure, and add special expression to call Syntax-directed operation which requires AST Value and algorithm name.
We gather which AST can be applied to a function. Also, to use child node of AST, we passes child node of AST to function as argument.

% \begin{itemize}
%   \item \textbf{Global Algorithms} Global algorithms are possible to be accessed
%     by any other algorithms. The top-level algorithm \( \code{RunJobs} \) is also global.
%   \item \textbf{Type-dependent Algorithms} Several algorithms have different
%     bodies dependent on types of their owner data. In ECMAScript specifications,
%     there are two different data types; language types and specification types.
%     The language types are explicitly used in JavaScript languages such as
%     \( \code{Boolean}, \code{Number}, \code{String} \), and \( \code{Object} \).
%     However, specification types are only used in specifications such as
%     Lists, Records, and Lexical Environments. Among them, some data types
%     have their own member algorithms. For example, language type \( \code{Object} \)
%     has several internal methods described in abstract algorithms. It depends
%     on 
% 
%     several data types have their own member algorithms. For example, 
%     
%     For example, the internal method
%     \( \code{[[Call]]} \) has different semantics dependent on its owner object.
%     Moreover, the owner objects are used in the body of algorithms as well.
%     Thus, we extend the parameters to accept its owner object as an implicit argument.
%   \item \textbf{Syntax-directed Algorithms} The semantics of each JavaScript syntax
%     are written in 
% \end{itemize}

% \subsection{semantic 컴파일하기 위해 하는 것 + specification 구조}
% 기본적인 원칙 :
% \begin{itemize} 
% \item 자바스크립트로 알고리즘 이름이랑 파라미터, 스텝을 가져와서 코어의 글로벌 ID에 매치 시켜줌
% \item initial Heap을 생성해줌
% \end{itemize}
% 매치를 시킨다는 뜻은 코어의 글로벌 ID에 코어 함수 형태로 알고리즘을 넣는 것임
%

>>>>

\section{Implementation}\label{sec:impl}

In this section, we demonstrate the implementation detail of \( \tool \).
We target to extract JavaScript syntax and semantics from
ECMAScript 2020 specification. It is the next ECMAScript version
that is actively updated in GitHub repository~\cite{es2020}
to be released in the next year. Moreover, Ecma Technical Committee 39 (TC39)
officially provides the conformance test suite, test262, for ECMAScript 2020.

\subsection{Excluded Language Features}
We decide to not support several language features in ECMAScript 2020:

\paragraph{No-Strict Mode}
Since ECMAScript 5, JavaScript introduced strict mode to \textit{intentianally} provide
slightly different semantics. It provides more stable semantics by preventing
unsafe actions or throwing originally ignored errors.
However, ECMAScript specification says:
``Strict mode does not restrict or modify any aspect of the ECMAScript semantics
that must operate consistently across multiple source text units.''
Thus, we decide to only support strict mode JavaScript to reduce the labors.

\paragraph{Intricate Built-ins}
JavaScript provides various built-in object to support helper functions in standard.
However, several built-in libraries are orthogonal but labor-intensive
to be manually modeled. Thus, we decide to not support
four orgothonoal libraries; \( \code{Math} \), \( \code{Date} \),
\( \code{RegExp} \), and \( \code{JSON} \).

\paragraph{Modules}
Moreover, we also do not support modules introduced since ES6 (ECMAScript 2015).
With the \( \code{import} \) keyword, a JavaScript program
imports other JavaScript components defined with the \( \code{export} \)
keyword in modules. In order to support modules in semantics, we should
model the functionality to communicate with file system to access
JavaScript module files. However, modules do not critically affect
the overall semantics. Thus, we decide to not support modules.

\subsection{Comple Rules for ECMAScript}

\begin{table}
  \centering
  \[
    \begin{array}{c|r|r|r|r|r|r|r}
      \text{name}
      & \multicolumn{1}{c|}{\code{Stmt}}
      & \multicolumn{1}{c|}{\code{Expr}}
      & \multicolumn{1}{c|}{\code{Cond}}
      & \multicolumn{1}{c|}{\code{Value}}
      & \multicolumn{1}{c|}{\code{Ty}}
      & \multicolumn{1}{c|}{\code{Ref}}
      & \multicolumn{1}{c}{\code{SecNo}}\\\hline
      \text{\# rules}
      & \inred{XX}
      & \inred{XX}
      & \inred{XX}
      & \inred{XX}
      & \inred{XX}
      & \inred{XX}
      & \inred{XX}\\\hline
    \end{array}
  \]
  \caption{The compile rules for ECMAScript specifications.}
  \label{table:rules}
\end{table}

The \textsf{Algorithm Compiler} requires compile rules to compile
given abstract algorithms into \( \ires \) functions.
We define compiler rules for ECMAScript specifications
and Table~\ref{table:rules} describes them.
The statement compile rule \( \code{Stmt} \) contains
\inred{XX} compile rules and generates \( \ires \) instructions.
The \( \code{Expr} \), \( \code{Cond} \), and \( \code{Value} \)
compile rules denote expressions, conditions, and values, respectively.
While all of them generates \( \ires \) expressions,
we discriminate them because they have different use cases.
The \( \code{Ty} \) denotes type names and generates
string primitives in \( \ires \). The \( \code{Ref} \)
represents the references including identifiers lookups or
member accesses of map-like structures. It generates
\( \ires \) reference expressions.
Finally, the Section number \( \code{SecNo} \) compile rule
represents sections numbers such as \( \code{9.4.1.1} \),
and it converts such numbers into the corresponding name strings
such as \( \code{"BoundFunctionObject.Call"} \).

\subsection{Global Setting}

We manually implemented \textsf{Global Setting} for ECMAScript 2020.
It consists of the structure of built-in objects and ECMAScript data types.

\subsubsection{ECMAScript Data Types}

In ECMAScript, the Section \textsf{6 ECMAScript Data Types and Values} describes
data types used in ECMAScript specifications. They not only represent
data types explicitely exposed into JavaScript evaluations, but also
represent all internal data types used in the specification.
For example, the Environement Records are record types from identifiers
into bindings for values. They are not explicitely accessed in JavaScript
programs but defined internally in the specification.

Each ECMAScript data type consists of several fields to store relevant values.
Thus, we should implement such structures to correctly evaluate JavaScript
programs. Moreover, some fields of ECMAScript data types might be
abstract algorithms. We call them as \textit{method-like abstract algorithms}
and they have implicitly get reciever object as an argument.
In order to mimic such implicit behaviors, we added a special variable
\( \code{this} \) as the first parameter. And we modify algorithm compiler
to detect cases and pass the receiver object into the method-like abstract
algorithms. For example, an Environement Record has the \textsf{DeleteBinding (N)}
abstract algorithm as one of its fields. Then, it has two parameters the special
parameter \( \code{this} \) and normal one \( \code{N} \).
The following algorithm step is compiled into the \( \ires \) instruction
\( \code{return DclRec.DeleteBinding(DclRec, N)} \).
\begin{center}
  \includegraphics[width=0.3\textwidth]{img/method-like-example.png}
\end{center}

\subsubsection{Built-in Objects}
In ECMAScript, built-in objects are pre-defined functions with several
built-in functions. For example, the array object \( \code{Array} \)
is the constructor of array objects, and its prototype property
\( \code{Array.prototype} \) has built-in functions for array objects.
For example, \( \code{[1,2,3].flat()} \) calls the
\( \code{Array.prototype.flat} \) built-in function with the array
\( \code{[1,2,3]} \). Built-in functions are also abstract algorithms
thus each of them is automatically converted into an \( \ires \) function.
However, the structured of built-in objects should be implemented manually.
Thus, we implement built-in objects in Scala and connect their properties
with the extracted \( \ires \) functions.

Some built-in objects are named as intrinsic objects when they are
are explicitly referenced in abstrat algorithms.
Intrinsic objects have their own names as aliases in the table
of the Section \textsf{6.1.7.4 Well-known Intrinsic Objects}.
Thus, we extract such alias information into \textsf{Global Setting}
to utilize it during evaluations.
